\documentclass[8pt, letterpaper, oneside]{extarticle}
\usepackage{fullpage, setspace, comment, url, verbatim, graphicx}
\usepackage[top=0.5in, bottom=0.5in, left=0.5in, right=0.5in]{geometry} 
\usepackage[pdfborder="0 0 0"]{hyperref}
\usepackage{enumitem}

\setlength{\parskip}{1ex plus 0.2ex minus 0.2ex}
\setlength{\parindent}{0pt}
\setcounter{secnumdepth}{6}
\linespread{1.1}

\setlist[enumerate,1]{label=(\alph*)}
\setlist[enumerate,2]{label=(\roman*)}

\newcounter{para}
\newcommand\numparagraph{\par\refstepcounter{para}\textbf{\thepara}\hspace{5mm}}

\title{Residential Tenancy Agreement and Lease}
\date{}

\begin{document}

 \maketitle

 \numparagraph{\textbf{INTRODUCTION:}\hspace{2mm}} \textbf{\#\#FIXME\#\#:landlord-name} (``Owner'') rents to \textbf{\#\#FIXME\#\#:tenant-name} (``Tenant'') and Tenant agrees to rent \textbf{\#\#FIXME\#\#:address}, \#\#FIXME\#\#:location-description (the ``Premises''). No other portion of the building (the ``Building''), where the Premises is located, is included for lease unless expressly provided for in this Residential Tenancy Agreement and Lease (the ``Agreement''). The Premises is provided furnished (see attached furniture inventory). The appliances provided at inception of the tenancy are described in the furniture inventory.

 \numparagraph{\textbf{TERM:}\hspace{2mm}} The term of this rental shall begin on \textbf{\#\#FIXME\#\#:start-date} and end on \textbf{\#\#FIXME\#\#:end-date}, and thereafter shall be month-to-month on the same terms and conditions as stated herein, save any changes lawfully made until terminated.

 \numparagraph{\textbf{PHYSICAL POSSESSION:}\hspace{2mm}} If owner is unable to deliver possession of the Premises at the commencement of the term, Owner shall not be liable for any damage caused thereby, nor shall this Agreement be void or voidable, but Tenant shall not be liable for any rent until possession is delivered.
 
 \numparagraph{\textbf{RENT:}\hspace{2mm}} The initial monthly base rent for the Premises shall be US \textbf{\#\#FIXME\#\#:rent-amount}. All rent is due and payable in full in advance of the first day of each and every month (the ``Due Date'') without offsets, deductions or credits. All rent shall be payable to Owner or such other person as Owner shall designate in writing. Tenant agrees always to pay rent by personal check, cashier's check, money order, or other means specified in the \textbf{Acceptable Payment Methods} addendum and not use cash unless specifically requested by owner. In the even of roommates, or another form of multiple occupancy, Tenant understands and agrees that rent shall be paid with a single payment and that it is up to Tenant to collect individual funds independently in order to submit a combined, single payment. Nothing in this Paragraph shall be construed as a direction by Owner to make payment in a particular manner; instead it is a promise by Tenant to pay in the manner set forth herein. Tenant bears the risk of loss or delay of any payment made by mail. Owner must receive mailed rent payments on or before the Due Date. Rent for any partial month shall be pro-rated at the rate of one thirtieth of the monthly rant per day. Owner may apply any payment made by Tenant to any obligation of Tenant to Owner notwithstanding any dates or other direction from Tenant that accompanies such payment. Any attempt by Tenant to allocate a payment in any other way shall be null and void, including the use or application of a restrictive endorsement on the face of any check. Owner will accept rent payments only from the Actual Tenant(s). No third part checks will be accepted, nor shall Owner be liable to Tenant in any way as a result of refusing any third party check. Should owner elect to accept a third party check such acceptance shall not be construed as a waiver of this provision.

 \numparagraph{\textbf{SECURITY DEPOSIT:}\hspace{2mm}} Before the commencement of the term, Tenant shall pay a security deposit of US \textbf{\#\#FIXME\#\#:security-deposit-amount} (the ``Security Deposit'') for the purposes set for in Civil Code Section 1950.5. No trust relationship between Owner and Tenant is created because of the Security Deposit and Owner may commingle the Security Deposit with other funds of Owner. Owner may retain such amounts of the Security Deposit as allowed by law including, but not limited to, amounts required to remedy future defaults by Tenant in any obligation under this Agreement to restore, replace, repairs, or return personal property or appurtenances, exclusive of ordinary wear and tear. Owner shall, within the time period alloted by law, refund any balance after such deductions to Tenant, after Tenant has vacated the Premises. Tenant shall not be deemed to have vacated the Premises for purposes of this Paragraph until
  \begin{enumerate}
   \item Tenant returns to owner all keys to the Premises, and
   \item Tenant has surrendered the Premises to Owner free and empty of all persons claiming any right to possess the Premises.
   \end{enumerate}
 Any balance of the Security Deposit and an accounting of any deductions therefrom will be mailed to the Tenant at the Premises unless Tenant provides, in writing to Owner, a mailing address to which the balance, if any, of the Security Deposit and the accounting should be sent. Owner's check or other draft refunding any balance of the Security Deposit may be made in the name of any or all of the original tenants regardless of the party who in fact made the deposit and regardless of the identity of the persons then occupying the Premises. Tenant may not apply the Security Deposit or any portion thereof to the last month's rent.

 If required by law, Owner shall pay to Tenant simple interest as directed by such law, less deductions, on the amount held as Security Deposit, provided this tenancy does not terminate before the Security Deposit has been held for one (1) year. Said payment of interest shall be made once a year commencing with the date the Security Deposit has been held for a year. Upon Tenant's surrender of the Premises, if the security Deposit is insufficient to remedy Tenant's default in rent, to repair damages caused by tenant, or to clean the Premises, Owner may use form the accrued unpaid interest such amounts as are necessary for those purposes. Accrued unpaid interest or balance thereof, if any, shall be mailed to Tenant at Tenant's las known address in the same manner as any refund of the Security Deposit.

 Owner may increase the Security Deposit up to the maximum allowed by law at any time with notice. The parties agree that the Security Deposit is not rent and therefore not subject to any local rent control law.

 If Owner applies any portion of the Security Deposit to obligations of Tenant at any time during the tenancy, Tenant must, upon thirty (30) days written notice, reinstate the Security Deposit to its full original amount. Owner may apply the Security Deposit during the term of the tenancy for any purpose allowed by law, and in such case, upon (30) days' written notice to Tenant, Tenant shall restore the Security Deposit to the full amount provided herein.

 \numparagraph{\textbf{FAILURE TO PAY:}\hspace{2mm}} Pursuant to Civil Code Section 1785.26, Tenant is hereby notified that a negative credit report reflecting on Tenant's credit record may be submitted to a credit reporting agency if Tenant fails to fulfill the terms of their credit obligations, such as the financial obligations of this Agreement.
 
 \numparagraph{\textbf{LATE PAYMENTS:}\hspace{2mm}} Tenant and owner agree that Owner will sustain costs and damage as a result of any late payment of rent but that it will be impracticable or extremely difficult to fix the actual damage. Therefore, the following sum represents a reasonable and fair estimate by Owner and Tenant of the actual damage that would be sustained. Tenant agrees to pay a late charge equal to US \textbf{\$35} for any payment of rent not received by Owner with \textbf{seven (7) business days} of the Due Date. The provision for payment of a late charge does not constitute a grace period, and Owner may serve a \emph{Three-Day Notice to Pay Rent or Quit} on the day after the Due Date. Owner and Tenant agree that Tenant paying the rent late on three (3) separate occasions within any twelve (12) month period shall constitute habitual late payment of rent and may be considered a just cause for eviction. Payment of a late charge does not cure the late payment for purposes of establishing habitual late payment of rent. The late charge shall be imposed for failure to pay on portion of the rent, including those portions allocated to parking and storage.
 
 \numparagraph{\textbf{RETURNED CHECKS:}\hspace{2mm}} Tenant and Owner agree that Owner will sustain costs and damage as a result of a check which is not honored by the bank on which it is drawn, for any reason, but that it will be impracticable or extremely difficult to fix the actual damage. Therefore, the following sum represents a reasonable and fair estimate by Owner and Tenant of the actual damage that would be sustained. Tenant agrees to pay Owner the additional sum of US \textbf{\$35} as a reimbursement of the expenses incurred by owner. A dishonored check shall constitute late payment of rent and shall be subject to the provisions of Paragraph 7 above regarding late payment, including but not limited to habitual late payment of rent. Such charges shall be immediately due and payable upon notice to Tenant. Failure to immediately pay the charges shall constitute a default under the terms of this agreement.
 
 Owner reserves the right to demand payment of rent by certified funds, cashier's check or money order for all future payments in the event of any such returned check or any other monetary default by Tenant, and rent tendered in any other form bay be refused by Owner. Nothing in this Paragraph shall limit other remedies available to owner as a payee of a dishonored check. Owner and Tenant agree that three (3) returned check in any twelve (12) month period shall constitute a frequent return of check due to insufficient funds and may be considered just cause for eviction.
 
 \numparagraph{\textbf{PARTIES TO AGREEMENT:}\hspace{2mm}} This Agreement is between owner and each named Tenant who is a signatory to this Agreement, individually and severally. Named signatory Tenants are jointly and severally responsible for the performance of their obligations under this Agreement, including the payment of rent until such time as the tenancy in its entirety is terminated and the Premises relinquished to owner, regardless of whether any named Tenant occupied the Premises.
 
 \numparagraph{\textbf{INDIVIDUAL LIABILITY:}\hspace{2mm}} Each person who signs this Agreement, whether or not said person is or remains in possession of the Premises, shall be jointly and severally responsible for the full performance of each and every obligation of this Agreement, including, but not limited to, the payment of all rent due and the payment of costs to remedy damages to the premises, regardless of whether such damages were caused by Tenant, Tenant's guests, or Tenant's invitees. This joint and several liability applies for as long as any one of the Tenants remains in possession.

 \numparagraph{\textbf{OCCUPANCY:}\hspace{2mm}} Tenant(s) named in paragraph 1 of this Agreement and no other is the only ``original occupant'' who took possession of the Premises pursuant to this Agreement. ``Original occupant'' can only be the person, or persons, who took occupancy of the Premises at the inception of the tenancy. All other persons who are not ``original occupants'' shall be considered ``subtenants''. Tenant may have overnight guests on the Premises for no more than fifteen (15) consecutive days or thirty (3) days in a calendar year, and no more than two (2) overnight guests per bedroom at any one time. Persons staying more than fifteen (15) consecutive days or more than thirty (3) days in any calendar year shall not be considered original occupants of the Premises. Tenant must obtain prior written approval of Owner if an invitee or overnight guess of Tenant will be present at the Premises for more than fifteen (15) consecutive days or thirty (30) days in a calendar year. Violation of the provisions of this section shall be deemed a substantial and material breach of this Agreement and is agreed to be a just cause for eviction

 \numparagraph{\textbf{INSPECTION OF PREMISES:}\hspace{2mm}} Tenant has been provided a Move-In/Move-Out Itemized Statement. Furnishings, equipment, plumbing, heating and electrical systems including smoke detectors, where applicable, are operative and are deemed satisfactory by Tenant unless Owner is notified in writing by Tenant to the contrary. The failure by Tenant to return to owner an acknowledged copy of the Move-In/Move-Out Itemized Statement within 48 hours after Tenant moves in shall be acknowledgement by Tenant that the Premises are habitable and in good condition.

 \numparagraph{\textbf{USE:}\hspace{2mm}} The Premises shall be used as a permanent, full-time dwelling for residential purposes only and for no other reason. No retail, commercial, or professional use of the Premises shall be made, unless such use conforms to applicable zoning laws and the prior written consent of the Owner is obtained in advance of such proposed use. As a condition for granting such permission, Owner may require that Tenant obtain liability insurance for the benefit of Owner. Tenants may not store any personal property outside their assigned dwelling until unless otherwise allowed by the terms herein. Owner has the right to remove any unlawfully stored items of personal property without notice.

 \numparagraph{\textbf{NUISANCE:}\hspace{2mm}} Tenant shall not commit, nor permit to be committed, any waste or nuisance upon, in, or about the Premises, nor shall Tenant create or permit a substantial interference with the comfort, safety, or enjoyment of Owner and/or other occupants of the Building or their guests or invitees.

 \numparagraph{\textbf{FINES AND PENALTIES:}\hspace{2mm}} Tenant is responsible for any fines or other costs occasioned by violations of the law by Tenant or Tenant's guests on the Premises or property while Tenant is in possession. If any such fines or costs are levied against Owner/Agent, Tenant agrees to pay such fines or costs attributed to Tenant's tenancy or the conduct of Tenant, Tenant's guests or others at the Premises, upon receipt of an invoice from Owner/Agent. The obligation to pay fines and costs assessed against Owner/Agent may be in addition to any assessed directly against Tenant.

 \numparagraph{\textbf{ASSIGNMENT AND SUBLETTING:}\hspace{2mm}} 
  \begin{enumerate}
   \item \textbf{No Assignment or Subletting: Except as Owner is required to permit by law, Tenant may not assign this Agreement or sublet the Premises or any portion of the Premises. This obligation of Tenant is intended as a strict and absolute prohibition against subletting and assignment and may not be waived by either party. If for any reason this prohibition is deemed to be invalid or unenforceable, then Paragraph 16 (b) shall apply.}
   
\vspace{5mm}\textbf{Tenant Initial Here: \underline{\hspace{3cm}}}

   \item Consent to Assignment or Subletting Required: Tenant agrees and covenants not to assign this Agreement or sublet the Premises or any portion of the Premises without first obtaining the prior written consent of Owner. This obligation may not be waived by either party.
%FIXME Section 14? Surely paragraph 16?

Only where Tenant requests consent to sublet or assign under Section 14(b) or whenever the law requires Owner to reasonably give consent to a request to sublet or assign (e.g., where the lease or rental agreement specifies the number of tenants to reside in a premises, or where the new occupant is the spouse or domestic partner, child, parent, grandchild, grandparent, siblings, or the spouse or domestic partner -- as defined in San Francisco Administrative Code Sections 62.1 through 62.8 -- of such relatives), the following provisions shall apply:
    \begin{enumerate}
     \item Any request to sublet must come in writing from Tenant before Subtenant takes occupancy of the Premises. The request must be given by certified mail or hand delivered to Owner at the place for giving notices as set forth in this Agreement
     \item Any request for subletting or assignment must be on the basis of a one-for-one replacement of the departing Tenant(s). Tenant may only request replacement of a departing Tenant or Tenants with an equal or fewer number of new tenants.
     \item Proposed Subtenant or Assignee, if requested by Owner, must complete Owner's standard application, or, in the event Owner fails to provide an application or has no standard application form, proposed Subtenant or Assignee must, upon request provide sufficient information to allow Owner to conduct a typical background check, including credit information, income information, references, and background information.
     \item Tenant must provide Owner five (5) business days to process proposed Subtenant's or Assignee's application
     \item Proposed new Subtenant or Assignee must meet the regular reasonable application standards of Owner.
     \item Tenant must not, without good cause, request Owner's consent to a new Subtenant or new Assignee more than one (1) time per existing Tenant residing in the Premises during the previous twelve (12) months.
     \end{enumerate}

\vspace{5mm}\textbf{Tenant Initial Here: \underline{\hspace{3cm}}}
   \end{enumerate}

 \numparagraph{\textbf{SMOKING:}\hspace{2mm}}  Smoking is not permitted in the Premises or in any other area of the Building. Tenant shall inform his or her guest(s) of this smoking prohibition. Tenant shall promptly notify Owner in writing of any incident where tobacco smoke is migrating into the Premises from sources outside of the Premises. Tenant acknowledges that Owner's adoption of this policy does not make Owner the guarantor of Tenant's heath of the smoke-free condition of the areas listed above. However, Owner shall take reasonable steps to enforce this provision. Owner shall not be required to take steps in response to smoking unless Owner has actual knowledge or has been provided with written notice. Owner and Tenant agree that the other Tenants of the Building are the third-party beneficiaries of this provision. A tenant may sue another Tenant to enforce this provision but does not have the right to evict another Tenant. Any lawsuit between Tenants regarding this provision shall not create a presumption that the Owner has breached this Agreement. A breach of this provision by the Tenant shall deemed a material breach of the Agreement and may be just cause for eviction.

 \numparagraph{\textbf{PETS:}\hspace{2mm}} No pets or animals are allowed in or about the Premises, even temporarily or with a visiting guest, except as allowed by law or by the express written consent of Owner, which may be unreasonably withheld. Any such consent is conditioned upon Tenant completing, signing, and returning to Owner the Pet Agreement, which shall become part of the Agreement.

 \numparagraph{\textbf{WATERBEDS:}\hspace{2mm}} Waterbeds and/or liquid-filled furniture are prohibited in accordance with Civil Code Section 1940.5. If the Premises are located in a structure for which the original Certificate of Occupancy was issued after January 1, 1973, then such furniture may be permitted on upon written consent of Owner, and upon the completion of a Waterbed Agreement, which shall become part of this Agreement.
 
 \numparagraph{\textbf{ROOF/FIRE ESCAPES:}\hspace{2mm}} Use of the roof and/or the fire escapes by Tenant, Tenant's guests, or Tenant's invitees is limited to emergency egress only. No other use is permitted, including but not limited to, the placement of personal property. No storage of any kind will be permitted on fire escapes or in other common areas. Owner reserves the right to remove any unauthorized personal property at any time without notice.

 \numparagraph{\textbf{STORAGE:}\hspace{2mm}} Storage is allowed pursuant to the attached Storage Agreement.

 \numparagraph{\textbf{PARKING:}\hspace{2mm}} This Agreement does not provide for parking of any motor vehicle or motorcycle anywhere in or about the Premises, the Building, and/or the driveway(s).

 \numparagraph{\textbf{UTILITIES:}\hspace{2mm}} Tenant shall pay directly for all utilities, services and charges provided to the Premises except for those listed as follows:
  \begin{enumerate}
   \item water,
   \item garbage,
   \item HOA dues,
   \item structure insurance, \&
   \item unit insurance.
   \end{enumerate}
 For utilities required to be paid directly by Tenant, Tenant must place all utilities in his or her name promptly. Tenant agrees to comply with any energy or water conservation, or utility-sharing programs implemented by Owner. Tenant understands that the rent paid by all Tenants is partially determined by the cost of utilities. Nothing contained herein prevents Owner from passing through to Tenant utility costs as provided by law. Tenant shall be provided access to the Building and the Premises for the installation of utility and communication lines and services as required by law and upon prior written consent by Owner. To the extent that any utilities, services and charges provided to the Premises are not paid by Tenant as set forth above, Tenant agrees to only use such utilities, services and charges provided to the premises which are reasonably necessary for the ordinary comfort and safety of Tenant to reside in the Premises. Excessive use of such utilities, services and charged provided to the Premises shall constitute a material breach of this Agreement. Tenant may not charge any device, including vehicles, in Building common areas or in designated parking/storage spaces without Owner's express written consent.
 
 \numparagraph{\textbf{INTERRUPTION OF SERVICES:}\hspace{2mm}} Owner shall not be liable to Tenant or to any other person for damages, nor shall Owner be in default under this Agreement, for any interruption or reduction of utilities or services caused by someone other than Owner, or by Owner due to circumstances beyond Owner's reasonable control.

 \numparagraph{\textbf{MAINTENANCE AND REPAIRS:}\hspace{2mm}} Tenant shall, at Tenant's expense, at all time maintain the Premises, furnishings and appliances, if any, in a clean and good condition and shall surrender the same upon termination of tenancy in the same condition as received (excepting normal wear and tear). Tenant understand that Tenant is responsible for the cost of repair of all damages in or about the Premises whether caused by Tenant, Tenant's guests, or Tenant's invitees.
 
 Tenant may not make any alterations to cable or telephone wiring without prior written consent of Owner or Owner's agent, except as provided for in the addendum entitled ``\textbf{Authorization to Change Telecommunications Providers}''. The consent request regarding proposed alterations to inside wiring shall include the name, address, and telephone number of any new telecommunications provider. Tenant shall hold owner harmless and indemnify Owner as to any mechanic's lien recordation or proceeding caused by Tenant. Tenant agrees to pay all costs resulting from the alteration and agrees to pay Owner any costs incurred as a result of restoring the inside wiring to the condition at the time of move-in, except for reasonable wear and tear.

 Except in an emergency, maintenance and repair requests must be made in writing and delivered to Owner or Owner's agent. Such notice shall also be deemed permission to enter the Premises to perform such maintenance or repairs in accordance with Civil Code Section 1954. Tenant may not place any unreasonable restrictions upon such access or entry. The Premises shall be presumed to be in a safe and habitable condition unless and until Tenant provides written notice to the contrary and said notice is received by Owner.

 In the event that the Premises contains landscaping under the exclusive control of Tenant, Tenant shall be required to properly maintain the landscaping at all times during the tenancy. The failure to maintain landscaping shall constitute a breach of a material lease covenant.

 Tenant acknowledges that the Premises and the Building from time to time may require renovations or repairs to keep them in good condition and repair and that such work may result in temporary loss of use of portions of the Building or the Premises and may inconvenience Tenant. Tenant agrees that any such loss shall not constitute a reduction in housing services, severance of housing service, or otherwise warrant a reduction in rent.

 \numparagraph{\textbf{ALTERATIONS:}\hspace{2mm}} Tenant shall not remodel renovate, paint, refinish floors, or otherwise alter the Premises, common areas, or any other parts of the Building. Tenant shall not apply adhesive paper to any cabinets, walls, or doors; nor shall Tenant hand any plants, planters or lighting fixtures from ceilings or walls; nor shall Tenant tack, nail or glue any coverings to floors or walls without prior written consent of Owner. Tenant shall not install or operate any washing machines, clothes dryers, portable dishwashers, deep-freeze units (or other such appliances), pianos, or outside antennae on the Premises without prior written consent of Owner. No plants, planters or plant boxes may be places directly on floors, carpets, window ledges or on fire escapes. Upon termination of tenancy, Owner shall have the option, at Owner's sole discretion, to require Tenant to restore the Premises to the original condition as received excepting normal wear and tear.

 \numparagraph{\textbf{SATELLITE DISHES:}\hspace{2mm}} Any satellite Dish installation shall be subject to all of the following rules and conditions:
  \begin{enumerate}
   \item Satellite Dish must be installed within the Premises or inside balcony railings or windows;
   \item Satellite Dish may not exceed one (1) meter in diameter;
   \item installation must comply with reasonable safety standards;
   \item installation must not damage Premises, Premises walls or other appurtenances;
   \item Tenant remains strictly liable for any injury or damage to persons or property caused by the Satellite Dish, and Tenant agrees to maintain sufficient liability coverage against any such injury or damage. Proof of such insurance must be provided to Owner, with Owner listed as an ``Additional Insured'', prior to approval of installation and upon each renewal of coverage.
   \end{enumerate}
 
 \numparagraph{\textbf{LOCKS:}\hspace{2mm}} Tenant shall not change any lock, alarm, or place additional locking devices upon any door or window of the Premises without the prior written consent of Owner. In the event of such installation, Tenant shall provide Owner with keys to such lock or device within 48 hours of Owner's request. Any expense incurred by Owner as a result of Tenant's action, such as changing of locks, shall be reimbursed by tenant upon demand. Once installed, an approved lock may not be removed even when the Premises is vacated. Keys to the Premises are the exclusive property of Owner. Tenant shall not consign keys to the Premises to any other person without the prior written consent of Owner. In the event that any keys to the Premises are lost, Tenant shall be liable for the entire cost of all key and lock replacement, at the discretion of Owner, as required for the security of the Premises, the Building, and the Building occupants. All keys must be returned to Owner when Tenant vacates. Tenant shall be charged for the cost of new locks and keys if all keys are not returned.
 
 \numparagraph{\textbf{DAMAGE TO PREMISES:}\hspace{2mm}} If the Premises are damaged by fire, flood, earthquake, or from any other cause so as to render them uninhabitable and therefore destroyed, the tenancy is terminated, unless restored pursuant to law.

 \numparagraph{\textbf{ENTRY AND INSPECTION:}\hspace{2mm}} Owner shall have the right to enter the Premises pursuant to California Civil Code Section 1945. Owner shall give Tenant reasonable notice of the intention of Owner/Owner's agent to enter the Premises and shall enter only during normal business hours; unless otherwise agreed by Tenant. For the purposes of this Paragraph, normal business hours shall be defined as 7:00 AM to 7:00 PM, every day of the week. Tenant may not place any unreasonable restrictions upon such entry. If, however, Owner reasonably believes that an emergency exists (such as fire or flood) which requires immediate entry, such entry may be made without prior notice to Tenant.
 
 If Tenant has, after written notice to cease, continued to deny Owner access to the Premises, as required by State law, such failure is a substantial breach of this Agreement and is a just cause for eviction.

 If the Premises or the Building is required by any government agency, lender, or insurer to undergo inspections, repairs or alterations, Tenant agrees to cooperate fully with Owner so that all such inspections, repairs or alterations are made in as expeditions and efficient a manner as possible.

 \numparagraph{\textbf{SMOKE DETECTION DEVICE:}\hspace{2mm}} The Premises are equipped with a functioning smoke detection device(s), and Tenant shall be responsible for testing the device as necessary and immediately reporting and problems, maintenance or need for repairs to Owner. If battery operated, Tenant is responsible for changing the detector's battery as necessary. Owner shall have a right to enter the Premises to check and maintain the smoke detection device as provided by law. It is expressly understood that Tenant may not, at any time, disable or remove an installed smoke detection device, and to do so shall be considered a material breach of this Agreement and will be a just cause for eviction.

 \numparagraph{\textbf{CARBON MONOXIDE DETECTION DEVICE:}\hspace{2mm}} If the Premises are equipped with a functioning carbon monoxide detection device(s), Tenant shall be responsible for testing the device as necessary and immediately reporting and problems, maintenance or need for repairs to Owner. If battery operated, Tenant is responsible for changing the detector's battery as necessary. Owner shall have a right to enter the Premises to check and maintain the carbon monoxide detection device as provided by law. It is expressly understood that Tenant may not, at any time, disable or remove an installed carbon monoxide detection device, and to do so shall be considered a material breach of this Agreement and will be a just cause for eviction.

 \numparagraph{\textbf{LEAD DISCLOSURE:}\hspace{2mm}} Many homes and apartments built before 1978 have paint that contains lead (called lead-based paint). Lead from paint chips and dust can pose serious health hazards if not taken care of properly.

 \numparagraph{\textbf{ASBESTOS:}\hspace{2mm}} The Premises may contain asbestos or have original construction materials that contain asbestos.
 
 Damaging or disturbing the surface of asbestos-containing materials (ACMs) may increase the risk of exposure. Therefore, Tenant and Tenant's guests, contractors, or invitees shall not allow any action which may, in any way, disturb ACMs or any part of the Premises that may contain asbestos or ACMs. Tenant shall notify Owner immediately if Tenant knows or suspects that an ACM has been disturbed or if Tenant becomes aware of any ACM that is showing signs of deterioration.

 \numparagraph{\textbf{MOLD/MILDEW:}\hspace{2mm}} Tenant agrees to maintain the Premises in a manner that prevents the occurrence and infestation of mold or mildew in the Premises, including the use of bathroom exhaust fans and/or opening windows as necessary to avoid moisture build-up. Tenant agrees to uphold this responsibility in part by complying with the list of responsibilities in the addendum entitled ``\textbf{Mold Notification}''. By signing this Agreement Tenant represents and agrees that Owner has provided Tenant with such information, including but not limited to \emph{Addendum -- ``Mold Notification''}.

 \numparagraph{\textbf{HAZARDOUS MATERIALS DISCLOSURE:}\hspace{2mm}} Pursuant to the regulations of Proposition 65, enacted by the voters of California, Owner hereby makes the following required disclosure:
  \begin{quote}
   ``Warning - The Premises contains chemicals known to the State of California to cause cancer and birth defects or other reproductive harm.''
   \end{quote}

 \numparagraph{\textbf{HOLD HARMLESS:}\hspace{2mm}} Owner shall not be liable for any damages or injury to Tenant, or any other person, or to any property, occurring on the Premises or any part thereof, or in common areas thereof, unless such damage is the proximate result of the intentional or unlawful act of Owner, Owner's agents, or Owner's employees. Tenant shall indemnify, defend and hold owner and Owner's agents harmless from all claims of loss or damage to property and of injury to or death of any person or persons caused by the intentional acts or negligence of Tenant, Tenant's guests, licensees, or invitees occurring in or about the Premises including other areas of the Building, adjacent sidewalks, streets, etc. Tenant hereby expressly releases Owner and/or Owner's Agents from any and all liability for loss or damage to Tenant's property or effects whether in the Premises, garage, storerooms or any other location in or about about the Premises, arising out of any cause whatsoever, including but not limited to rain, plumbing leakage, fire or theft, except in the case that such damage has been adjudged to be the result of the gross negligence of Owner, Owner's employees, heirs, successors, assignees, and/or agents.

 \numparagraph{\textbf{ESTOPPEL CERTIFICATES:}\hspace{2mm}} Tenant(s), and all persons holding under them, shall within five (5) days after receipt from Owner, sign and return to Owner a certificate, statement or other such document signed by all occupants providing such information as owner may reasonably request, under penalty of perjury, including, but not limited to, the amount of base rent currently paid, the names of each occupant and their respective initial dates of occupancy, the date of the last rent increase, the status of each occupant entitled to special benefits based on age, health, disability, income, or other criteria under any provision of the local rent laws or other applicable law, the identity of furniture of fixtures that belong to Tenant, whether the rent includes any parking space or storage space, and the amount of any security deposit or prepaid rent and whether interest on said deposit has been paid and through which date. In addition, Tenant shall disclose, upon request, any information that Tenant believes would prevent any purchase of the Premises or would prevent Owner of a potential owner from moving into the Premises. Failure to deliver the above described document within the five (5) days shall be a material breach of this Agreement.

 \numparagraph{\textbf{INSURANCE / RENTER'S INSURANCE:}\hspace{2mm}} Generally, except under special circumstances, OWNER IS NOT legally responsible for loss to the Tenant's personal property, possessions or personal liability, and OWNER'S INSURANCE WILL NOT COVER such losses or damages In addition, if damage to Owner's property or an injury is caused by Tenant, Tenant's guest(s), invitees or child (children), Owner's insurance company may have the right to attempt to recover from Tenant(s) payments made under Owner's policy.
 
 Tenant is required to maintain renter's insurance on a policy form commonly know as HO-4 throughout the duration of the tenancy that includes at a minimum:
  \begin{enumerate}
   \item Personal Liability coverage of at lease US \$300,000.00 combined single limit per occurrence for Bodily Injury and Property Damage;
   \item the Premises listed above must be listed on the policy as the ``insured premises'' of the Tenant insured;
   \item Owner is listed as a Certificate Holder and, if acceptable to the insurance company, as an additional insured; and
   \item the carrier must provide thirty (30) day's notice of cancellation or non-renewal to Owner, except for non-payment of premium for which ten (10) days' notice shall be given.
   \end{enumerate}
 Owner strongly recommends that Tenant consult with a qualified insurance agent and obtain appropriate coverage for fire, theft, liability, and other perils.

 Tenant must provide proof of such insurance to Owner within thirty (30) days of the inception of the tenancy. In addition, Tenant must provide to owner proof on insurance on an annual basis. The failure to abide by this covenant shall constitute a material breach of this Agreement and is a just cause for eviction.

 \numparagraph{\textbf{NO WAIVER:}\hspace{2mm}} No failure of Owner to enforce any term of the Agreement will be deemed a waiver of that term or of any other term of the Agreement. The waiver by Owner of any term of the Agreement will not be deemed to be a waiver of any subsequent breach of the same or any other term of the Agreement, nor will any custom or practice which may develop between the parties be construed to waive or lessen the right of Owner to insist upon performance by Tenant of all the provisions of the Agreement, or support a claim of detrimental reliance by Tenant. Owner's acceptance of a partial payment of rent will not constitute a waiver of Owner's right to the full amount due, nor will Owner's acceptance of rent paid late ever constitute a waiver of Owner's right to evict Tenant for habitual late payment of rent. \textbf{This covenant of the Agreement cannot be waived by Owner.}

 \numparagraph{\textbf{APPLICATION:}\hspace{2mm}} Any Rental Application or related form submitted by Tenant is incorporated herein as though set forth in full. Any misrepresentations contained therein shall be considered a material incurable breach of the Agreement and may be just cause for eviction.

 \numparagraph{\textbf{HOUSE RULES:}\hspace{2mm}} Tenant agrees to abide by any and all house rules, whether made known before or after the date of the Agreement, including, but not limited to, rules with respect to noise, odors, disposal of refuse and use of common areas. Tenant has read, understands and agrees to be bound by the existing House Rules attached to and made part of the Agreement.

 \numparagraph{\textbf{CAPTIONS:}\hspace{2mm}} The captions in the Agreement are to assist the parties in reading the Agreement and are not a part of its term or provisions.

 \numparagraph{\textbf{INITIALS:}\hspace{2mm}} The Agreement is binding and enforceable even if a page of the Agreement is not initialed by either party or both.

 \numparagraph{\textbf{INVALIDITY:}\hspace{2mm}} The invalidity of partial invalidity of any provision of the Agreement shall not render the remainder of the Agreement invalid or unenforceable. Violation by Tenant of any applicable ordinance or statute shall be deemed sufficient cause for termination of tenancy. Notwithstanding any other provision of the Agreement, each and every expressed term and condition is deemed material by the parties.

 \numparagraph{\textbf{CONDOMINIUM REQUIREMENTS COMPLIANCE:}\hspace{2mm}} Tenant has been advised that the Premises is a condominium unit. Tenant understand and acknowledges that Tenant's occupancy of the Premises is to some extent governed by a Declaration of Covenants, Conditions \& Restrictions (CC\&Rs) and Rules \& Regulations of a Homeowners Association (``HOA''). All such documents are referred to as ``Governing Documents''. Tenant acknowledges that Tenant has been provided with the Governing Documents, or has had the opportunity to examine them, prior to entering into the Agreement, and that Tenant and Tenant's guests, invitees, and/or sublessees (if any) will at all times comply with the requirements of the Governing Documents, and that Tenant's failure or that of Tenant's guests, invitees, and/or sublessees to do so will constitute a material breach of the Agreement. Tenant further agrees that should Owner be assessed any fines or penalties by the HOA as a result of any conduct by Tenant or the parties associated with Tenant, Tenant will immediately pay to Owner, on Owner's demand, the full amount of any such penalty or assessment. Owner shall have no obligation to defend against the assessment with the HOA.

 \numparagraph{\textbf{ITEMS OF PERSONAL PROPERTY LEFT BEHIND:}\hspace{2mm}} Owner and Tenant hereby agree that any items of personal possessions left in the Premises after Tenant vacates, or is evicted, shall be considered refuse and of no value. Tenant hereby stipulates that any items of personal property left at the Premises after termination of the tenancy may be immediately discarded by owner without an auction, and that Owner may immediately discard Tenant's personal property without storing said property for any length of time.

 \numparagraph{\textbf{TERMINATION:}\hspace{2mm}} If Tenant intends to vacate at the end of the original term of this Agreement, of for any other reason after the original term of this Agreement, Tenant must give owner at least thirty (30) day prior, written notice of Tenant's intention to terminate the tenancy and vacate the premises. After the expiration of the original term of this Agreement, Owner may terminate the tenancy, in accordance with applicable law. Upon termination, Tenant shall completely vacate the Premises and any parking or storage areas; give written notice of Tenant's forwarding address; and deliver all keys, furnishings, if any, and the Premises to Owner in the same condition as received excepting normal wear and tear. Rent shall be due and payable through the end of the notice period or the date Tenant vacates, whichever comes later. Tenant may rescind said notice within five (5) calendar days after it is served on Owner without incurring liability to any person. Such rescission must be in writing, and delivered to Owner. Thereafter, if Tenant fails to vacate the Premises on or before the date set forth in Tenant's notice, Tenant shall be liable for any costs incurred by Owner or any third parties who relied upon Tenant's notice terminating the tenancy, and failure to vacate as agreed is a substantial violation of the terms of the tenancy and is a just cause for eviction. Tenant further agrees to defend, protect, indemnify and hold Owner harmless from any and all damages, lost rents, costs, expenses, losses, claims and liabilities, including attorney's fees, arising in any way out of Tenant's failure to comply with the provisions of Tenant's notice. Tenant's failure to pay any such sums within twenty (20) days after demand shall be deemed material breach of the Agreement. It is agreed by the parties to this Agreement that delivery of possession shall be deemed to occur when Tenant delivers the keys to the Premises to Owner or Owner's agent during normal business hours as stated in this Agreement.
 
 Any attempt by tenant to terminate the Agreement prior to the end of the original term shall be deemed to be a breach of the Agreement, and Owner shall be entitled to recover all damages occasioned thereby, including leasing commissions, advertising expenses, utilities maintained to show the Premises, and all unpaid rent until the term of the lease.

 \numparagraph{\textbf{ATTORNEY'S FEES:}\hspace{2mm}} If either party has to retain legal counsel to enforce any of the rights or obligations created under this Agreement, the prevailing party shall be entitled to recover from the non-prevailing party the prevailing party's reasonable attorney's fees and costs regardless of whether litigation is actually instituted.

 \numparagraph{\textbf{ENTIRE AGREEMENT:}\hspace{2mm}} This Agreement consists of the foregoing numbered paragraphs and attachments identified in this agreement herein:
  \begin{enumerate}
   \item House Rules
   \item Move-In/Move-Out Condition Report
   \item Mold Notification
   \item Disclosure of Information on Lead-Based Paint and Lead-Based Paint Hazards
   \item Pet Agreement
   \item Storage Agreement
   \item San Francisco Waste Disposal Addendum
   \item Acceptable Payment Methods Addendum
   \item Cleaning Obligations Addendum
   \item Authorization to Change Telecommunications Providers
   \item Transition Agreement
   \end{enumerate}
   
 The foregoing constitutes the entire agreement between the parties and may be modified only in writing signed by all parties except that Owner may change the terms of the tenancy and the Agreement pursuant to Civil Code Section 827. If Owner changes a term of the tenancy pursuant to Section 827, then by remaining in possession of the Premises when the change takes effect, Tenant is deemed by such affirmative act to have consented to the change.

 \vspace{5mm}

 \textbf{BY SIGNING BELOW TENANT(S) HEREBY ACKNOWLEDGE(S) READING AND UNDERSTANDING THE TERMS OF THIS AGREEMENT, AND FURTHER ACKNOWLEDGE(S) RECEIVING A COPY HEREOF:}

 \vspace{10mm}

 Tenant Signature: \hspace{5mm} \underline{\hspace{90mm}} \hspace{10mm} Date: \hspace{5mm} \underline{\hspace{40mm}}

 \vspace{10mm}

 Tenant Signature: \hspace{5mm} \underline{\hspace{90mm}} \hspace{10mm} Date: \hspace{5mm} \underline{\hspace{40mm}}

 \vspace{10mm}

 Tenant Signature: \hspace{5mm} \underline{\hspace{90mm}} \hspace{10mm} Date: \hspace{5mm} \underline{\hspace{40mm}}

 \vspace{10mm}

 \textbf{ACCEPTED AND AGREED TO BY OWNER}

 \vspace{10mm} 
 Owner Signature: \hspace{5mm} \underline{\hspace{90mm}} \hspace{10mm} Date: \hspace{5mm} \underline{\hspace{40mm}}

 \end{document}
